---
title: "Comparing economic conditions in Jersey City (for Shawn)"
author: "Andrew Ba Tran"
output:
  html_document:
  code_folding: hide
pdf_document: default
---
  
This is a more-focused geospatial analysis of unemployment rates in the census tracts chosen specifically by developers compared with the tracts they originally pitched to the state. 

The numbers are bucketed and colors are chosen randomly for so it would be easier for graphics to edit.

```{r setup, include=FALSE}
library(rgdal)
library(rgeos)
library(tidyverse)
library(tigris)
library(ggplot2)
library(geosphere)

library(tidycensus)
source("scripts/keys.R")

library(viridis)
library(ggmap)
library(DT)
library(leaflet)
options(tigris_use_cache = TRUE)

census_api_key(key)

library(readr)
library(lubridate)
library(dplyr)
library(stringr)
library(sp)
library(knitr)
```


```{r prepping_centroids, fig.width=9, fig.height=9, warning=F, message=F}
nj_h <- tracts("NJ", county="Hudson", cb=F)
nj_h_centroids <-  SpatialPointsDataFrame(gCentroid(nj_h, byid=TRUE), 
                                          nj_h@data, match.ID=FALSE)

nj_h_centroids <- as.data.frame(nj_h_centroids)
nj_h_centroids <- select(nj_h_centroids, GEOID, x, y)

nj_h_centroids$distance_proj1 <- 0
nj_h_centroids$distance_proj2 <- 0

for (i in 1:nrow(nj_h_centroids)) {
  nj_h_centroids$distance_proj1[i] <- distm (c(-74.063644, 40.734330), c(nj_h_centroids$x[i], nj_h_centroids$y[i]), fun = distHaversine)[,1]/ 1609
  nj_h_centroids$distance_proj2[i] <- distm (c(-74.03577, 40.72), c(nj_h_centroids$x[i], nj_h_centroids$y[i]), fun = distHaversine)[,1]/ 1609
}
```


# 65 Bay Street (2015)

```{r map2, fig.width=9, fig.height=6, warning=F, message=F}

# GEOIDs: 34017004400, 34017004500, 34017004600, 34017005200, 34017005300, 34017005500, 34017005600, 34017005801, 34017006000, 34017006100, 34017006200, 34017006400, 34017006500, 34017006700, 34017006800, 34017007600 


#GEOIDS: 3,4,5,6,7,8,11,12.01, 12.02, 15, 22, 23,25,26,30,31,32,33,44,45,46,50,52,53,55

hudson_unemployment_2012 <- get_acs(geography="tract", endyear=2015, variables= c("B23025_005E", "B23025_002E"), county = "Hudson", state="NJ")
hudson_unemployment_2012$moe <- NULL

hudson_unemployment_2012 <- spread(hudson_unemployment_2012,variable, estimate )
hudson_unemployment_2012$per_un <- round(hudson_unemployment_2012[,4]/hudson_unemployment_2012[,3]*100,2)
hudson_unemployment_2012 <- filter(hudson_unemployment_2012, GEOID!="34017006900" & GEOID!="34017980100")

proj1 <- c("34017004400", "34017004500", "34017004600", "34017005200", "34017005300", "34017005500", "34017005600", "34017005801", "34017006000", "34017006100", "34017006200", "34017006400", "34017006500", "34017006700", "34017006800", "34017007600")

hudson_unemployment_2012_sm <- filter(hudson_unemployment_2012, GEOID %in% proj1)

colnames(hudson_unemployment_2012_sm) <- c("GEOID", "name", "total", "unemployed", "unemp_rate")
hudson_unemployment_2012_sm$un_rate <- hudson_unemployment_2012_sm$unemp_rate$B23025_005
hudson_unemployment_2012_sm$unemp_rate <- NULL

hudson_unemployment_2012_sm$radius <- "gerrymandered"

proj1_half <- c("34017000300", "34017000400", "34017000500", "34017000600", "34017000700", "34017000800", "34017001100", "34017001201", "34017001202", "34017001500", "34017002200", "34017002300", "34017002500", "34017002600", "34017003000", "34017003100", "34017003200",
                "34017003300", "34017004400", "34017004500", "34017004600", "34017005000",
                "34017005200", "34017005300", "34017005500")

hudson_unemployment_2012_sm_half <- filter(hudson_unemployment_2012, GEOID %in% proj1_half)
colnames(hudson_unemployment_2012_sm_half) <- c("GEOID", "name", "total", "unemployed", "unemp_rate")
hudson_unemployment_2012_sm_half$un_rate <- hudson_unemployment_2012_sm_half$unemp_rate$B23025_005
hudson_unemployment_2012_sm_half$unemp_rate <- NULL

hudson_unemployment_2012_sm_half$radius <- "adjacent tracts"

hudson_unemployment_sm <- rbind(hudson_unemployment_2012_sm, hudson_unemployment_2012_sm_half)

hudson_unemployment_sm$name <- gsub(",.*", "", hudson_unemployment_sm$name)

nj_hf <- fortify(nj_h, region="GEOID")


nj_hud <- left_join(nj_hf, hudson_unemployment_sm, by=c("id"="GEOID"))
nj_hud <- filter(nj_hud, !is.na(radius))

nj_hud$nj_mid <- cut(nj_hud$un_rate, 6)
colors <- colorRampPalette(c("white", "red"))(length(levels(nj_hud$nj_mid)))

nj_map <- ggplot()
#nj_map <- nj_map + geom_polygon(data=nj_hf, aes(x=long, y=lat, group=group), fill=NA, color="black", size=.1)
nj_map <- nj_map + geom_polygon(data=nj_hud, aes(x=long, y=lat, group=group, fill=factor(nj_mid)), color="black", size=.5)
nj_map <- nj_map + facet_wrap(~radius)
nj_map <- nj_map + coord_map() 
#nj_map <- nj_map + scale_x_discrete()
nj_map <- nj_map + scale_fill_manual(drop=FALSE, values=c("gray", "yellow", "orange", "red", "green", "purple"), na.value="#EEEEEE", name="Unemployment rate")
#nj_map <- nj_map + scale_fill_viridis(option = "heat", direction=-1, name = "Unemployment rate")
#nj_map <- nj_map + scale_color_viridis(option = "heat", direction=-1)
nj_map <- nj_map + theme_nothing(legend=TRUE) 
nj_map <- nj_map + labs(x=NULL, y=NULL, title="65 Bay Street (2015)")
nj_map <- nj_map + theme(panel.grid.major = element_line(colour = NA))
nj_map <- nj_map + theme(text = element_text(size=15))
nj_map <- nj_map + theme(plot.title=element_text(face="bold", hjust=.4))
nj_map <- nj_map + theme(plot.subtitle=element_text(face="italic", size=9, margin=margin(l=20)))
nj_map <- nj_map + theme(plot.caption=element_text(size=12, margin=margin(t=12), color="#7a7d7e", hjust=0))
nj_map <- nj_map + theme(legend.key.size = unit(1, "cm"))
nj_map <- nj_map + annotate("segment", x = -74.03577, xend = -74.005, y = 40.72, yend = 40.72, colour = "tomato", size=.5) 
nj_map <- nj_map + annotate("point", x = -74.03577, y = 40.72, colour = "lightblue", size = 1) 
nj_map <- nj_map + annotate("text", x = -73.98217, y = 40.72, label = "65 Bay Street", size=3, colour="gray30") 


print(nj_map)
#ggsave(nj_map, device="pdf", "map2a.pdf")

nj_map <- ggplot()
nj_map <- nj_map + geom_polygon(data=nj_hf, aes(x=long, y=lat, group=group), fill=NA, color="black", size=.1)
nj_map <- nj_map + geom_polygon(data=nj_hud, aes(x=long, y=lat, group=group, fill=factor(nj_mid)), color="black", size=.5)
nj_map <- nj_map + facet_wrap(~radius)
nj_map <- nj_map + coord_map() 
#nj_map <- nj_map + scale_x_discrete()
nj_map <- nj_map + scale_fill_manual(drop=FALSE, values=c("gray", "yellow", "orange", "red", "green", "purple"), na.value="#EEEEEE", name="Unemployment rate")
#nj_map <- nj_map + scale_fill_viridis(option = "heat", direction=-1, name = "Unemployment rate")
#nj_map <- nj_map + scale_color_viridis(option = "heat", direction=-1)
nj_map <- nj_map + theme_nothing(legend=TRUE) 
nj_map <- nj_map + labs(x=NULL, y=NULL, title="65 Bay Street (2015)")
nj_map <- nj_map + theme(panel.grid.major = element_line(colour = NA))
nj_map <- nj_map + theme(text = element_text(size=15))
nj_map <- nj_map + theme(plot.title=element_text(face="bold", hjust=.4))
nj_map <- nj_map + theme(plot.subtitle=element_text(face="italic", size=9, margin=margin(l=20)))
nj_map <- nj_map + theme(plot.caption=element_text(size=12, margin=margin(t=12), color="#7a7d7e", hjust=0))
nj_map <- nj_map + theme(legend.key.size = unit(1, "cm"))
nj_map <- nj_map + annotate("segment", x = -74.03577, xend = -74.005, y = 40.72, yend = 40.72, colour = "tomato", size=.5) 
nj_map <- nj_map + annotate("point", x = -74.03577, y = 40.72, colour = "lightblue", size = 1) 
nj_map <- nj_map + annotate("text", x = -73.98217, y = 40.72, label = "65 Bay Street", size=3, colour="gray30") 


print(nj_map)
#ggsave(nj_map, device="pdf", "map2b.pdf")

```

For this project, officials said the qualifying rate using 2015 annual average data is 9.3 percent (1.5 times the United States not seasonally adjusted unemployment rate for the same time period).

So this project would've qualified based on only the gerrymandered version since unemployment was too low in the other ways of grouping the census tracts.


```{r table2, fig.height=6, fig.width=9, warning=F, message=F}

sm_table <- hudson_unemployment_sm %>%
  group_by(radius) %>%
  summarize(average_unemployment=round(mean(un_rate, na.rm=T),2), median_unemployment=round(median(un_rate, na.rm=T),2)) %>%
  arrange(average_unemployment)

kable(sm_table)


```

