---
title: "Profiles on Jersey City project tracts"
author: "Andrew Ba Tran"
output:
  html_document:
    code_folding: hide
---

This geospatial analysis of Jersey City census tracts in Hudson county was used to assist the reporting of the story: **[Jared Kushner and his partners used a program meant for job-starved areas to build a luxury skyscraper](https://www.washingtonpost.com/investigations/jared-kushner-and-his-partners-used-a-program-meant-for-job-starved-areas-to-build-a-luxury-skyscraper/2017/05/31/9c81b52c-4225-11e7-9869-bac8b446820a_story.html)**

Two projects from Jared Kushner's development company received millions of federal funding designed to benefit poor areas. To qualify for low-cost financing, the Kushner group daisy-chained the low-unemployment rate area around the tower site to neighborhoods with poor neighborhoods almost four miles away.

This analysis looks at census tracts that Kushner's development group said would benefit from their project in their proposal. It compares the 5-year ACS data used in their methodology with the most recent year available (2015). 

```{r setup, include=FALSE}
# I use the development version of ggplot2
# install.packages("devtools")
# devtools::install_github("tidyverse/ggplot2")

library(tidycensus)
library(tidyverse)
source("../scripts/keys.R")
source("../scripts/theme_nothing.R")
library(tigris)
#library(sf)
library(viridis)
library(ggmap)
library(DT)
library(leaflet)
options(tigris_use_cache = TRUE)

census_api_key(key)

library(readr)
library(lubridate)
library(dplyr)
library(stringr)
library(sp)



```

# 1 Journal Square (2017)

**County:** Hudson

**Census tracts:** 19, 46, 53, 66, 67, 71

**Address:** 1 Journal Square, Jersey City, NJ 07306

The developer's methodology can be revealed by clicking the 'code' button on the right.

```{r methodology1, fig.width=9, fig.height=6, message=F, warning=F}

#* A ratio is developed for the census tracts based on the ACS 2009 - 2013 Census data and then applied to the 2015 annual average data for the derivation county

#* The ACS 2009 - 2013 Census employment and unemployment figures for the tract are divided by the corresponding ACS 2009 - 2013 Census employment and unemployment data for the derivation county

#* The resulting ratios are thena pplied to the 2015 annual average employment and unemployment data for the derivation county

#* These employment and unemployment figures are then summed to determine the labor force figure

#* Finally, the unemployment rate is calcuated by dividing the unemployment figure by the labor force figure

#* The qualifying rate using 2015 annual average data is 8.0 percent (1.5 times the US not seasonally adjusted unemployment rate for the same time period)
```

This analysis seeks to replicate their process as closely as possible.

```{r data_prep, warning=F, message=F}
# GEOIDs: 34017001900, 34017004600, 34017005300, 34017006600, 34017006700, 34017007100

hudson_unemployment_2013 <- get_acs(geography="tract", endyear=2013, variables= c("B23025_005E", "B23025_002E"), county = "Hudson", state="NJ")
hudson_unemployment_2015 <- get_acs(geography="tract", endyear=2015, variables= c("B23025_005E", "B23025_002E"), county = "Hudson", state="NJ")

hudson_unemployment_2013$moe <- NULL
hudson_unemployment_2015$moe <- NULL

hudson_unemployment_2013 <- spread(hudson_unemployment_2013,variable, estimate )
hudson_unemployment_2015 <- spread(hudson_unemployment_2015,variable, estimate )
hudson_unemployment_2013$per_un <- round(hudson_unemployment_2013[,4]/hudson_unemployment_2013[,3]*100,2)
hudson_unemployment_2015$per_un <- round(hudson_unemployment_2015[,4]/hudson_unemployment_2015[,3]*100,2)

hudson_unemployment_2013 <- filter(hudson_unemployment_2013, GEOID!="34017006900" & GEOID!="34017980100")
hudson_unemployment_2015 <- filter(hudson_unemployment_2015, GEOID!="34017006900" & GEOID!="34017980100")

hudson_unemployment_2013$rank <- rank(hudson_unemployment_2013$per_un)
hudson_unemployment_2015$rank <- rank(hudson_unemployment_2015$per_un)

proj1 <- c("34017001900", "34017004600", "34017005300", "34017006600", "34017006700", "34017007100")

hudson_unemployment_2013_sm <- filter(hudson_unemployment_2013, GEOID %in% proj1)
hudson_unemployment_2015_sm <- filter(hudson_unemployment_2015, GEOID %in% proj1)

colnames(hudson_unemployment_2013_sm) <- c("GEOID", "name", "total", "unemployed", "unemp_rate_2013", "2013_rank")
colnames(hudson_unemployment_2015_sm) <- c("GEOID", "name", "total", "unemployed", "unemp_rate_2015", "2015_rank")

hudson_unemployment_2013_sm <- select(hudson_unemployment_2013_sm, id=GEOID, name, unemp_rate_2013, `2013_rank`)
hudson_unemployment_2015_sm <- select(hudson_unemployment_2015_sm, id=GEOID, unemp_rate_2015, `2015_rank`)
hudson_unemp <- left_join(hudson_unemployment_2013_sm, hudson_unemployment_2015_sm)
hudson_unemp$diff <- hudson_unemp$unemp_rate_2015$B23025_005 - hudson_unemp$unemp_rate_2013$B23025_005

avg_2013 <- mean(hudson_unemp$unemp_rate_2013$B23025_005)
avg_2015 <- mean(hudson_unemp$unemp_rate_2015$B23025_005)
avg_diff <- avg_2015 - avg_2013

hudson_unemp$ur_2013 <- hudson_unemp$unemp_rate_2013$B23025_005 
hudson_unemp$ur_2015 <- hudson_unemp$unemp_rate_2015$B23025_005 

hudson_unemp$unemp_rate_2013 <- NULL
hudson_unemp$unemp_rate_2015 <- NULL

hudson_unemp2 <- select(hudson_unemp, name, id, ur_2013, ur_2015)
hudson_unemp2 <- gather(hudson_unemp2, "year", "unemployment_rate", 3:4)
hudson_unemp3 <- select(hudson_unemp, tract=name, `unemployment rate 2013`=ur_2013, `unemployment rate 2015`=ur_2015, `2013 rank`=`2013_rank`, `2015 rank`=`2015_rank`)
hudson_unemp3$tract <- gsub(",.*", "", hudson_unemp3$tract)
datatable(hudson_unemp3)
```

There are `r nrow(hudson_unemployment_2013)` census tracts within Hudson County. 

The tracts associated with this project include the areas with the highest and lowest unemployment rates in the county. The tract where 1 Journal Square was constructed tied for 62nd place in lowest unemployment rate at 8.6 percent in 2013. Tract 67 was included by planners within the project borders but is nearly two and a half miles away from the site and had an unemployment rate of 24 percent— the second worst rate in the county.

According to the application, the combined unemployment rate is greater than or equal to the qualifying rate of 8 percent (150 percent of the US rate for the same time period).

According to our calculations (and setting aside margin of error for now), the average unemployment rate in these tracts between 2007 and 2013 was estmated to be `r round(avg_2013,1)` percent. Between 2013 and 2015, that rate changed to `r round(avg_2015,1)` percent. 

**That's a difference of `r round(avg_diff,1)`.**

```{r double_check, fig.width=9, fig.height=6, message=F, warning=F}
# bringing in jersey city tract boundaries

nj_h <- tracts("NJ", county="Hudson", cb=F)
nj_hf <- fortify(nj_h, region="GEOID")

nj_hud <- left_join(nj_hf, hudson_unemp2)
nj_hud <- filter(nj_hud, !is.na(year))
nj_hud$year <- gsub("ur_", "", nj_hud$year)
one_journal_square <- c("40.734330, -74.063644")

nj_map <- ggplot()
nj_map <- nj_map + geom_polygon(data=nj_hf, aes(x=long, y=lat, group=group), fill=NA, color="black", size=.1)
nj_map <- nj_map + geom_polygon(data=nj_hud, aes(x=long, y=lat, group=group, fill=unemployment_rate), color="black", size=.5)
nj_map <- nj_map + facet_wrap(~year)
nj_map <- nj_map + coord_map() 
nj_map <- nj_map + scale_fill_viridis(option = "heat", direction=-1, name = "Unemployment rate")
nj_map <- nj_map + scale_color_viridis(option = "heat", direction=-1)
nj_map <- nj_map + theme_nothing(legend=TRUE) 
nj_map <- nj_map + labs(x=NULL, y=NULL, title="1 Journal Square (2017)")
nj_map <- nj_map + theme(panel.grid.major = element_line(colour = NA))
nj_map <- nj_map + theme(text = element_text(size=15))
nj_map <- nj_map + theme(plot.title=element_text(face="bold", hjust=.4))
nj_map <- nj_map + theme(plot.subtitle=element_text(face="italic", size=9, margin=margin(l=20)))
nj_map <- nj_map + theme(plot.caption=element_text(size=12, margin=margin(t=12), color="#7a7d7e", hjust=0))
nj_map <- nj_map + theme(legend.key.size = unit(1, "cm"))
nj_map <- nj_map + annotate("segment", x = -74.063644, xend = -74.035, y = 40.734330, yend = 40.734330, colour = "tomato", size=.5) 
nj_map <- nj_map + annotate("point", x = -74.063644, y = 40.734330, colour = "lightblue", size = 1) 
nj_map <- nj_map + annotate("text", x = -74.01, y = 40.734330, label = "1 Journal Square", size=3, colour="gray30") 

print(nj_map)

## slope graph

hudson_unemp2$year <- gsub("ur_", "", hudson_unemp2$year)
hudson_unemp2$name <- gsub(",.*", "", hudson_unemp2$name)
hudson_unemp2_proj <- filter(hudson_unemp2, id=="34017001900")

sp <- ggplot(hudson_unemp2) 
sp <- sp + geom_line(aes(x=as.factor(year), y=unemployment_rate, group=id), size=1)
sp <- sp + geom_point(aes(x=as.factor(year), y=unemployment_rate), size=3)
sp <- sp + geom_line(data=hudson_unemp2_proj, aes(x=as.factor(year), y=unemployment_rate, group=id), size=1, color="tomato")
sp <- sp + geom_point(data=hudson_unemp2_proj, aes(x=as.factor(year), y=unemployment_rate), size=3, color="tomato")
sp <- sp + theme_minimal()
sp <- sp + scale_y_log10()
sp <- sp +geom_text(data = subset(hudson_unemp2, year=="2013"), 
                    aes(x = as.factor(year), y = unemployment_rate, label = paste0(name, ": ", unemployment_rate, "%")), 
                    size = 3, hjust = 1.7)
sp <- sp +  geom_text(data = subset(hudson_unemp2, year=="2015"), 
                      aes(x = as.factor(year), y = unemployment_rate, label =  paste0(unemployment_rate, "%")), 
                      size = 3, hjust = -1.5) 


sp <- sp+ theme(legend.position = "none", 
                panel.grid.major.y = element_blank(),
                panel.grid.minor.y = element_blank(),
                panel.grid.major.x = element_blank(), 
                axis.ticks.y = element_blank(),
                axis.ticks.x = element_blank(), 
                axis.title.y = element_blank(), 
                axis.text.y = element_blank(), 
                plot.title = element_text(size = 18))  
sp <- sp + ggtitle("Unemployment rate in tracts associated with 1 Journal Square (2017)")
sp <- sp + xlab("")
print(sp)
```

```{r map1, fig.width=9, fig.height=6, warning=F, message=F}
hudson_unemp4 <- subset(hudson_unemp2, year=="2013")
nj_merged <- geo_join(nj_h, hudson_unemp4, "GEOID", "id")

#nj_merged <- filter(nj_merged, !is.na(unemployment_rate))
nj_merged <- nj_merged[!is.na(nj_merged$unemployment_rate),]
pal_sb <- colorNumeric("Greens", domain=hudson_unemp2$unemployment_rate)
popup_sb <- paste0(nj_merged$name, ": ", as.character(nj_merged$unemployment_rate))


leaflet() %>%
  addProviderTiles("CartoDB.Positron") %>%
  setView(-74.063644, 40.734330, zoom = 13) %>% 
  addPolygons(data = nj_merged , 
              fillColor = ~pal_sb(nj_merged$unemployment_rate), 
              color = "#444444",
              fillOpacity = 0.7, 
              weight = 1, 
              smoothFactor = 0.2,
              highlight = highlightOptions(
                weight = 5,
                color = "#666",
                dashArray = "",
                fillOpacity = 0.7,
                bringToFront = TRUE),
              label=popup_sb,
              labelOptions = labelOptions(
                style = list("font-weight" = "normal", padding = "3px 8px"),
                textsize = "15px",
                direction = "auto")) %>%
  addMarkers(lng=-74.063644, lat=40.734330, popup="1 Journal Square") %>%
  addLegend(pal = pal_sb, 
            values = nj_merged$unemployment_rate, 
            position = "bottomright", 
            title = "Unemployment rate<br />2013")
```

----

# 1 65 Bay Street (2015)

**County:** Hudson

**Census tracts:** 44, 45, 46, 52, 53, 55, 56, 58.01, 60, 61, 62, 64, 65, 67, 68, 76

**Address:** 65 Bay Street, Jersey City, NJ

The developer's methodology can be revealed by clicking the 'code' button on the right.

```{r methodology3, fig.width=9, fig.height=6, message=F, warning=F}

#* A ratio is developed for the census tracts based on the ACS 2008 - 2012 Census data and then applied to 2014 annual average data fro the derivation county

#* The ACS 2008 - 2012 Census employment and unemployment figures for the tract are divided by the corresponding ACS 2008 - 2012 Census employment and unemployment data for the derivation county

#* The resulting ratios are then applied to the 2014 annual average employment and  unemployment data for the3 derivation county

#* These employment and unemployoment figures are then summed to determine the labor force figure

#* Finally, the unemployment rate is calculated by dividing the unemployment figure by the labor force figure

#* The qualifying rate using 2014 annual average data is 9.3 percent (1.5 times the US not seasonally adjusted unemployment rate for the same time period)
```

```{r data_prep3, warning=F, message=F}
# GEOIDs: 34017004400, 34017004500, 34017004600, 34017005200, 34017005300, 34017005500, 34017005600, 34017005801, 34017006000, 34017006100, 34017006200, 34017006400, 34017006500, 34017006700, 34017006800, 34017007600 

hudson_unemployment_2012 <- get_acs(geography="tract", endyear=2012, variables= c("B23025_005E", "B23025_002E"), county = "Hudson", state="NJ")
hudson_unemployment_2015 <- get_acs(geography="tract", endyear=2015, variables= c("B23025_005E", "B23025_002E"), county = "Hudson", state="NJ")

hudson_unemployment_2012$moe <- NULL
hudson_unemployment_2015$moe <- NULL

hudson_unemployment_2012 <- spread(hudson_unemployment_2012,variable, estimate )
hudson_unemployment_2015 <- spread(hudson_unemployment_2015,variable, estimate )
hudson_unemployment_2012$per_un <- round(hudson_unemployment_2012[,4]/hudson_unemployment_2012[,3]*100,2)
hudson_unemployment_2015$per_un <- round(hudson_unemployment_2015[,4]/hudson_unemployment_2015[,3]*100,2)

hudson_unemployment_2012 <- filter(hudson_unemployment_2012, GEOID!="34017006900" & GEOID!="34017980100")
hudson_unemployment_2015 <- filter(hudson_unemployment_2015, GEOID!="34017006900" & GEOID!="34017980100")

hudson_unemployment_2012$rank <- rank(hudson_unemployment_2012$per_un)
hudson_unemployment_2015$rank <- rank(hudson_unemployment_2015$per_un)

proj1 <- c("34017004400", "34017004500", "34017004600", "34017005200", "34017005300", "34017005500", "34017005600", "34017005801", "34017006000", "34017006100", "34017006200", "34017006400", "34017006500", "34017006700", "34017006800", "34017007600")

hudson_unemployment_2012_sm <- filter(hudson_unemployment_2012, GEOID %in% proj1)
hudson_unemployment_2015_sm <- filter(hudson_unemployment_2015, GEOID %in% proj1)

colnames(hudson_unemployment_2012_sm) <- c("GEOID", "name", "total", "unemployed", "unemp_rate_2012", "2012_rank")
colnames(hudson_unemployment_2015_sm) <- c("GEOID", "name", "total", "unemployed", "unemp_rate_2015", "2015_rank")

hudson_unemployment_2012_sm <- select(hudson_unemployment_2012_sm, id=GEOID, name, unemp_rate_2012, `2012_rank`)
hudson_unemployment_2015_sm <- select(hudson_unemployment_2015_sm, id=GEOID, unemp_rate_2015, `2015_rank`)
hudson_unemp <- left_join(hudson_unemployment_2012_sm, hudson_unemployment_2015_sm)
hudson_unemp$diff <- hudson_unemp$unemp_rate_2015$B23025_005 - hudson_unemp$unemp_rate_2012$B23025_005

avg_2012 <- mean(hudson_unemp$unemp_rate_2012$B23025_005)
avg_2015 <- mean(hudson_unemp$unemp_rate_2015$B23025_005)
avg_diff <- avg_2015 - avg_2012

hudson_unemp$ur_2012 <- hudson_unemp$unemp_rate_2012$B23025_005 
hudson_unemp$ur_2015 <- hudson_unemp$unemp_rate_2015$B23025_005 

hudson_unemp$unemp_rate_2012 <- NULL
hudson_unemp$unemp_rate_2015 <- NULL

hudson_unemp2 <- select(hudson_unemp, name, id, ur_2012, ur_2015)
hudson_unemp2 <- gather(hudson_unemp2, "year", "unemployment_rate", 3:4)
hudson_unemp3 <- select(hudson_unemp, tract=name, `unemployment rate 2012`=ur_2012, `unemployment rate 2015`=ur_2015, `2012 rank`=`2012_rank`, `2015 rank`=`2015_rank`)
hudson_unemp3$tract <- gsub(",.*", "", hudson_unemp3$tract)
datatable(hudson_unemp3)
```

There are 164 census tracts within Hudson County. 

The tracts associated with this project include the areas with the highest and lowest unemployment rates in the county. The tract where 65 Bay Street was constructed tied for 28th place in lowest unemployment rate at 5 percent in 2012. Tract 53 was included by planners within the project borders but is nearly three away from the site and had an unemployment rate of 24 percent— the worst rate in Hudson county that year.

According to the application, the combined unemployment rate is greater than or equal to the qualifying rate of 9.8 percent (150 percent of the US rate for the same time period).


According to our calculations (and ignoring margin of error for now), the average unemployment rate in these tracts between 2008 and 2012 was estmated to be `r round(avg_2012,1)` percent. Between 2012 and 2015, that rate changed to `r round(avg_2015,1)` percent. That's a difference of `r round(avg_diff,1)`. 


```{r double_check3, fig.width=9, fig.height=6, message=F, warning=F}
# bringing in jersey city tract boundaries

nj_h <- tracts("NJ", county="Hudson", cb=F)
nj_hf <- fortify(nj_h, region="GEOID")

nj_hud <- left_join(nj_hf, hudson_unemp2)
nj_hud <- filter(nj_hud, !is.na(year))
nj_hud$year <- gsub("ur_", "", nj_hud$year)

nj_map <- ggplot()
nj_map <- nj_map + geom_polygon(data=nj_hf, aes(x=long, y=lat, group=group), fill=NA, color="black", size=.1)
nj_map <- nj_map + geom_polygon(data=nj_hud, aes(x=long, y=lat, group=group, fill=unemployment_rate), color="black", size=.5)
nj_map <- nj_map + facet_wrap(~year)
nj_map <- nj_map + coord_map() 
nj_map <- nj_map + scale_fill_viridis(option = "heat", direction=-1, name = "Unemployment rate")
nj_map <- nj_map + scale_color_viridis(option = "heat", direction=-1)
nj_map <- nj_map + theme_nothing(legend=TRUE) 
nj_map <- nj_map + labs(x=NULL, y=NULL, title="65 Bay Street (2015)")
nj_map <- nj_map + theme(panel.grid.major = element_line(colour = NA))
nj_map <- nj_map + theme(text = element_text(size=15))
nj_map <- nj_map + theme(plot.title=element_text(face="bold", hjust=.4))
nj_map <- nj_map + theme(plot.subtitle=element_text(face="italic", size=9, margin=margin(l=20)))
nj_map <- nj_map + theme(plot.caption=element_text(size=12, margin=margin(t=12), color="#7a7d7e", hjust=0))
nj_map <- nj_map + theme(legend.key.size = unit(1, "cm"))
nj_map <- nj_map + annotate("segment", x = -74.03577, xend = -74.005, y = 40.72, yend = 40.72, colour = "tomato", size=.5) 
nj_map <- nj_map + annotate("point", x = -74.03577, y = 40.72, colour = "lightblue", size = 1) 
nj_map <- nj_map + annotate("text", x = -73.98217, y = 40.72, label = "65 Bay Street", size=3, colour="gray30") 

print(nj_map)

## slope graph

hudson_unemp2$year <- gsub("ur_", "", hudson_unemp2$year)
hudson_unemp2$name <- gsub(",.*", "", hudson_unemp2$name)
hudson_unemp2_proj <- filter(hudson_unemp2, id=="34017007600")

sp <- ggplot(hudson_unemp2) 
sp <- sp + geom_line(aes(x=as.factor(year), y=unemployment_rate, group=id), size=1)
sp <- sp + geom_point(aes(x=as.factor(year), y=unemployment_rate), size=3)
sp <- sp + geom_line(data=hudson_unemp2_proj, aes(x=as.factor(year), y=unemployment_rate, group=id), size=1, color="tomato")
sp <- sp + geom_point(data=hudson_unemp2_proj, aes(x=as.factor(year), y=unemployment_rate), size=3, color="tomato")
sp <- sp + theme_minimal()
sp <- sp + scale_y_log10()
sp <- sp +geom_text(data = subset(hudson_unemp2, year=="2012"), 
                    aes(x = as.factor(year), y = unemployment_rate, label = paste0(name, ": ", unemployment_rate, "%")), 
                    size = 3, hjust = 1.7)
sp <- sp +  geom_text(data = subset(hudson_unemp2, year=="2015"), 
                      aes(x = as.factor(year), y = unemployment_rate, label =  paste0(unemployment_rate, "%")), 
                      size = 3, hjust = -1.5) 


sp <- sp+ theme(legend.position = "none", 
                panel.grid.major.y = element_blank(),
                panel.grid.minor.y = element_blank(),
                panel.grid.major.x = element_blank(), 
                axis.ticks.y = element_blank(),
                axis.ticks.x = element_blank(), 
                axis.title.y = element_blank(), 
                axis.text.y = element_blank(), 
                plot.title = element_text(size = 18))  
sp <- sp + ggtitle("Unemployment rate in tracts associated with 65 Bay Street (2015)")
sp <- sp + xlab("")
print(sp)
```


```{r map3, fig.width=9, fig.height=6, warning=F, message=F}
hudson_unemp4 <- subset(hudson_unemp2, year=="2012")
nj_merged <- geo_join(nj_h, hudson_unemp4, "GEOID", "id")

#nj_merged <- filter(nj_merged, !is.na(unemployment_rate))
nj_merged <- nj_merged[!is.na(nj_merged$unemployment_rate),]
pal_sb <- colorNumeric("Greens", domain=hudson_unemp2$unemployment_rate)
popup_sb <- paste0(nj_merged$name, ": ", as.character(nj_merged$unemployment_rate))


leaflet() %>%
  addProviderTiles("CartoDB.Positron") %>%
  setView(-74.03577, 40.72,, zoom = 13) %>% 
  addPolygons(data = nj_merged , 
              fillColor = ~pal_sb(nj_merged$unemployment_rate), 
              color = "#444444",
              fillOpacity = 0.7, 
              weight = 1, 
              smoothFactor = 0.2,
              highlight = highlightOptions(
                weight = 5,
                color = "#666",
                dashArray = "",
                fillOpacity = 0.7,
                bringToFront = TRUE),
              label=popup_sb,
              labelOptions = labelOptions(
                style = list("font-weight" = "normal", padding = "3px 8px"),
                textsize = "15px",
                direction = "auto")) %>%
  addMarkers(lng=-74.03577, lat=40.72, popup="65 Bay Street") %>%
  addLegend(pal = pal_sb, 
            values = nj_merged$unemployment_rate, 
            position = "bottomright", 
            title = "Unemployment rate<br />2012")
```

# Calls to police by census tract for 65 Bay Street

```{r calls_and_shots, fig.height=6, fig.width=9, warning=F, message=F}
police_calls <- read_csv("../data/jcpd-all.csv")
shootings <- read_csv("../data/jerseycity2015shootings-use.csv")


shootings$datetime <- mdy_hm(shootings$datetime)


police_calls$datetime <- mdy_hm(police_calls$`Time Received`)

police_calls$blah <- ifelse(is.na(police_calls$datetime),police_calls$`Time Received`, "" )

police_calls$datetime2 <- mdy_hms(police_calls$blah)

police_calls$datetime3 <- ifelse(is.na(police_calls$datetime),as.character(police_calls$datetime2), as.character(police_calls$datetime))
police_calls$datetime3 <- ymd_hms(police_calls$datetime3)
police_calls$year <- year(police_calls$datetime3)
police_calls_2015 <- subset(police_calls, year==2015)

police_calls_2015 <- police_calls_2015 [!duplicated(police_calls_2015$`Event Number`),]


calls <- c("BURG ALARM COMMERCIAL PROP", "ASSAULT NO WEAPON", "USE/SALE OF DRUGS", "THEFT PROP FROM VEHICLE", "MOTOR VEHICLE THEFT", "THEFT FROM PERSON",
           "ROB COMERCIAL", "FOUND CDS/PARAPHERNALIA", "HOMICIDE")

police_calls_2015_sub <- police_calls_2015 %>%
  filter(str_detect(`Call Code Description`, "BURG ALARM COMMERCIAL PROP|ASSAULT NO WEAPON|USE/SALE OF DRUGS|THEFT PROP FROM VEHICLE|MOTOR VEHICLE THEFT|THEFT FROM PERSON|ROB COMERCIAL|FOUND CDS/PARAPHERNALIA|HOMICIDE"))

police_calls_2015_sub$type <- gsub(";.*", "", police_calls_2015_sub$`Call Code Description`)
police_calls_2015_sub <- select(police_calls_2015_sub, type, date=datetime3, lat=LATITUDE, lon=LONGITUDE)
shootings_2015_sub <- shootings
shootings_2015_sub$shooting_type <- paste("Shooting:", shootings_2015_sub$shooting_type)
shootings_2015_sub <- select(shootings, type=shooting_type, date=datetime, lat=latitude, lon=longitude)

sub_2015 <- rbind(police_calls_2015_sub, shootings_2015_sub)

coords <- sub_2015[c("lon", "lat")]

sp <- SpatialPoints(coords)
#plot(nj_h)
#plot(sp, col="red", add=TRUE)

proj4string(sp) <- "+proj=longlat +datum=NAD83 +no_defs +ellps=GRS80 +towgs84=0,0,0"
#proj4string(sp)

by_tract <- over(sp, nj_h)

sub_2015 <- cbind(sub_2015, by_tract)

tract_totals <- sub_2015 %>%
  group_by(GEOID, NAMELSAD, type) %>%
  summarize(total=n())

tract_totals <- tract_totals %>%
  group_by(type) %>%
  mutate(percent=round(total/sum(total, na.rm=T)*100,2))

tract_totals$type <- ifelse(tract_totals$type=="Fatal", "Shooting: Fatal", tract_totals$type)
tract_totals$type <- ifelse(tract_totals$type=="Non-fatal", "Shooting: Non-fatal", tract_totals$type)

tract_totals <- filter(tract_totals, !is.na(GEOID))

proj1 <- c("34017004400", "34017004500", "34017004600", "34017005200", "34017005300", "34017005500", "34017005600", "34017005801", "34017006000", "34017006100", "34017006200", "34017006400", "34017006500", "34017006700", "34017006800", "34017007600")

tract_totals_a  <- filter(tract_totals, GEOID %in% proj1)

shootings_tracts <- filter(tract_totals_a, type=="Shooting: Fatal" | type=="Shooting: Non-fatal")
tract_totals_a <- filter(tract_totals_a, type!="Shooting: Fatal" & type!="Shooting: Non-fatal")

nj_h <- tracts("NJ", county="Hudson", cb=F)
nj_hf <- fortify(nj_h, region="GEOID")


nj_tracts <- left_join(nj_hf, tract_totals_a, by=c("id"="GEOID"))

#nj_tracts <- filter(nj_tracts, !is.na(year))
nj_tracts <- nj_tracts[!is.na(nj_tracts$percent),]

#nj_tracts$year <- gsub("ur_", "", nj_hud$year)

nj_map <- ggplot()
#nj_map <- nj_map + geom_polygon(data=nj_hf, aes(x=long, y=lat, group=group), fill=NA, color="black", size=.1)
nj_map <- nj_map + geom_polygon(data=nj_tracts, aes(x=long, y=lat, group=group, fill=percent), color="black", size=.5)
nj_map <- nj_map + facet_wrap(~type, ncol=2)
nj_map <- nj_map + coord_map() 
nj_map <- nj_map + scale_fill_viridis(option = "heat", direction=-1, name = "Percent of crime calls")
nj_map <- nj_map + scale_color_viridis(option = "heat", direction=-1)
nj_map <- nj_map + theme_nothing(legend=TRUE) 
nj_map <- nj_map + labs(x=NULL, y=NULL, title="Percent of crime calls in 2015")
nj_map <- nj_map + theme(panel.grid.major = element_line(colour = NA))
nj_map <- nj_map + theme(text = element_text(size=15))
nj_map <- nj_map + theme(plot.title=element_text(face="bold", hjust=.4))
nj_map <- nj_map + theme(plot.subtitle=element_text(face="italic", size=9, margin=margin(l=20)))
nj_map <- nj_map + theme(plot.caption=element_text(size=12, margin=margin(t=12), color="#7a7d7e", hjust=0))
nj_map <- nj_map + theme(legend.key.size = unit(1, "cm"))
nj_map <- nj_map + annotate("segment", x = -74.03577, xend = -74.005, y = 40.72, yend = 40.72, colour = "tomato", size=.5) 
nj_map <- nj_map + annotate("point", x = -74.03577, y = 40.72, colour = "lightblue", size = 1) 
nj_map <- nj_map + annotate("text", x = -73.99217, y = 40.72, label = "65 Bay Street", size=3, colour="gray30") 

print(nj_map)
datatable(tract_totals_a[,2:5])


```

```{r shootings, fig.width=9, fig.height=6, warning=F, message=F}


nj_tracts <- left_join(nj_hf, shootings_tracts, by=c("id"="GEOID"))

#nj_tracts <- filter(nj_tracts, !is.na(year))
nj_tracts <- nj_tracts[!is.na(nj_tracts$percent),]

#nj_tracts$year <- gsub("ur_", "", nj_hud$year)

nj_map <- ggplot()
#nj_map <- nj_map + geom_polygon(data=nj_hf, aes(x=long, y=lat, group=group), fill=NA, color="black", size=.1)
nj_map <- nj_map + geom_polygon(data=nj_tracts, aes(x=long, y=lat, group=group, fill=percent), color="black", size=.5)
nj_map <- nj_map + facet_wrap(~type, ncol=2)
nj_map <- nj_map + coord_map() 
nj_map <- nj_map + scale_fill_viridis(option = "heat", direction=-1, name = "Percent of shootings")
nj_map <- nj_map + scale_color_viridis(option = "heat", direction=-1)
nj_map <- nj_map + theme_nothing(legend=TRUE) 
nj_map <- nj_map + labs(x=NULL, y=NULL, title="Shootings in 2015")
nj_map <- nj_map + theme(panel.grid.major = element_line(colour = NA))
nj_map <- nj_map + theme(text = element_text(size=15))
nj_map <- nj_map + theme(plot.title=element_text(face="bold", hjust=.4))
nj_map <- nj_map + theme(plot.subtitle=element_text(face="italic", size=9, margin=margin(l=20)))
nj_map <- nj_map + theme(plot.caption=element_text(size=12, margin=margin(t=12), color="#7a7d7e", hjust=0))
nj_map <- nj_map + theme(legend.key.size = unit(1, "cm"))
nj_map <- nj_map + annotate("segment", x = -74.03577, xend = -74.005, y = 40.72, yend = 40.72, colour = "tomato", size=.5) 
nj_map <- nj_map + annotate("point", x = -74.03577, y = 40.72, colour = "lightblue", size = 1) 
nj_map <- nj_map + annotate("text", x = -73.99217, y = 40.72, label = "65 Bay Street", size=3, colour="gray30") 

print(nj_map)
datatable(shootings_tracts[,2:5])

```

```{r one_more_map, fig.width=9, fig.height=6, warning=F, message=F}

cof <- colorFactor(c("red", "purple"), domain=c("Non-fatal", "Fatal"))
# mapping based on type

leaflet() %>%
  addProviderTiles("CartoDB.Positron") %>%
  setView(-74.03577, 40.72, zoom = 13) %>% 
  addPolygons(data = nj_merged,
              group="Unemployment",
              fillColor = ~pal_sb(nj_merged$unemployment_rate), 
              color = "#444444",
              fillOpacity = 0.4, 
              weight = 1, 
              smoothFactor = 0.2,
              highlight = highlightOptions(
                weight = 5,
                color = "#666",
                dashArray = "",
                fillOpacity = 0.7,
                bringToFront = TRUE),
              label=popup_sb,
              labelOptions = labelOptions(
                style = list("font-weight" = "normal", padding = "3px 8px"),
                textsize = "15px",
                direction = "auto")) %>%
  addMarkers(lng=-74.03577, lat=40.72, popup="65 Bay Street") %>%
  addLegend(pal = pal_sb, 
            values = nj_merged$unemployment_rate, 
            position = "bottomright", 
            title = "Unemployment rate<br />2012") %>% 
  addCircleMarkers(data=shootings, ~longitude, ~latitude, popup=shootings$location, weight = 3, radius=4, color=~cof(shooting_type), stroke = F, fillOpacity = 0.5, group="Shootings")  %>%
  addLegend("bottomright", colors= c("red", "purple"), labels=c("Fatal", "Non-fatal"), title="Shootings") %>%
    addLayersControl(
    overlayGroups = c("Unemployment", "Shootings"),
    options = layersControlOptions(collapsed = FALSE)
  )

```