---
title: "Comparing economic conditions in Jersey City (for graphics)"
author: "Andrew Ba Tran"
output:
  html_document:
    code_folding: hide
  pdf_document: default
---
  
This is a more-focused geospatial analysis of unemployment rates in the census tracts chosen specifically by developers compared with the tracts immediately adjacent to the project site. This is [another version](http://wpinvestigative.github.io/kushner_eb5_census/tracts.html) with tracts within a certain distance of the site.

The numbers are bucketed and colors are chosen randomly for so it would be easier for graphics to edit.

```{r setup, include=FALSE}
library(rgdal)
library(rgeos)
library(tidyverse)
library(tigris)
library(ggplot2)
library(geosphere)

library(tidycensus)
source("../scripts/keys.R")
library(viridis)
library(ggmap)
library(DT)
library(leaflet)
options(tigris_use_cache = TRUE)

census_api_key(key)

library(readr)
library(lubridate)
library(dplyr)
library(stringr)
library(sp)
library(knitr)
```


```{r prepping_centroids, fig.width=9, fig.height=9, warning=F, message=F}
nj_h <- tracts("NJ", county="Hudson", cb=F)
nj_h_centroids <-  SpatialPointsDataFrame(gCentroid(nj_h, byid=TRUE), 
                                          nj_h@data, match.ID=FALSE)

nj_h_centroids <- as.data.frame(nj_h_centroids)
nj_h_centroids <- select(nj_h_centroids, GEOID, x, y)

nj_h_centroids$distance_proj1 <- 0
nj_h_centroids$distance_proj2 <- 0

for (i in 1:nrow(nj_h_centroids)) {
  nj_h_centroids$distance_proj1[i] <- distm (c(-74.063644, 40.734330), c(nj_h_centroids$x[i], nj_h_centroids$y[i]), fun = distHaversine)[,1]/ 1609
  nj_h_centroids$distance_proj2[i] <- distm (c(-74.03577, 40.72), c(nj_h_centroids$x[i], nj_h_centroids$y[i]), fun = distHaversine)[,1]/ 1609
}
```

# 1 Journal Square (2017)

```{r map1, fig.width=9, fig.height=6, warning=F, message=F}

# GEOIDs: 34017001900, 34017004600, 34017005300, 34017006600, 34017006700, 34017007100
# GEIODs: 19, 20 18, 17, 9.02, 12.02, 71

hudson_unemployment_2013 <- get_acs(geography="tract", endyear=2013, variables= c("B23025_005E", "B23025_002E"), county = "Hudson", state="NJ")
hudson_unemployment_2013$moe <- NULL

hudson_unemployment_2013 <- spread(hudson_unemployment_2013,variable, estimate )
hudson_unemployment_2013$per_un <- round(hudson_unemployment_2013[,4]/hudson_unemployment_2013[,3]*100,2)
hudson_unemployment_2013 <- filter(hudson_unemployment_2013, GEOID!="34017006900" & GEOID!="34017980100")

proj1 <- c("34017001900", "34017004600", "34017005300", "34017006600", "34017006700", "34017007100")

hudson_unemployment_2013_sm <- filter(hudson_unemployment_2013, GEOID %in% proj1)

colnames(hudson_unemployment_2013_sm) <- c("GEOID", "name", "total", "unemployed", "unemp_rate")
hudson_unemployment_2013_sm$un_rate <- hudson_unemployment_2013_sm$unemp_rate$B23025_005
hudson_unemployment_2013_sm$unemp_rate <- NULL

hudson_unemployment_2013_sm$radius <- "gerrymandered"

proj1_half <- c("34017001900", "34017002000", "34017001800", "34017001700", "34017000902", "34017001202", "34017007100")

hudson_unemployment_2013_sm_half <- filter(hudson_unemployment_2013, GEOID %in% proj1_half)
colnames(hudson_unemployment_2013_sm_half) <- c("GEOID", "name", "total", "unemployed", "unemp_rate")
hudson_unemployment_2013_sm_half$un_rate <- hudson_unemployment_2013_sm_half$unemp_rate$B23025_005
hudson_unemployment_2013_sm_half$unemp_rate <- NULL

hudson_unemployment_2013_sm_half$radius <- "adjacent tracts"

hudson_unemployment_sm <- rbind(hudson_unemployment_2013_sm, hudson_unemployment_2013_sm_half)

hudson_unemployment_sm$name <- gsub(",.*", "", hudson_unemployment_sm$name)

nj_hf <- fortify(nj_h, region="GEOID")


nj_hud <- left_join(nj_hf, hudson_unemployment_sm, by=c("id"="GEOID"))
nj_hud <- filter(nj_hud, !is.na(radius))


nj_hud$nj_mid <- cut(nj_hud$un_rate, 6)
colors <- colorRampPalette(c("white", "red"))(length(levels(nj_hud$nj_mid)))

nj_map <- ggplot()
#nj_map <- nj_map + geom_polygon(data=nj_hf, aes(x=long, y=lat, group=group), fill=NA, color="black", size=.1)
nj_map <- nj_map + geom_polygon(data=nj_hud, aes(x=long, y=lat, group=group, fill=factor(nj_mid)), color="black", size=.5)
nj_map <- nj_map + facet_wrap(~radius)
nj_map <- nj_map + coord_map() 
#nj_map <- nj_map + scale_x_discrete()
nj_map <- nj_map + scale_fill_manual(drop=FALSE, values=c("gray", "yellow", "orange", "red", "green", "purple"), na.value="#EEEEEE", name="Unemployment rate")
nj_map <- nj_map + theme_nothing(legend=TRUE) 
nj_map <- nj_map + labs(x=NULL, y=NULL, title="1 Journal Square (2017)")
nj_map <- nj_map + theme(panel.grid.major = element_line(colour = NA))
nj_map <- nj_map + theme(text = element_text(size=15))
nj_map <- nj_map + theme(plot.title=element_text(face="bold", hjust=.4))
nj_map <- nj_map + theme(plot.subtitle=element_text(face="italic", size=9, margin=margin(l=20)))
nj_map <- nj_map + theme(plot.caption=element_text(size=12, margin=margin(t=12), color="#7a7d7e", hjust=0))
nj_map <- nj_map + theme(legend.key.size = unit(1, "cm"))
nj_map <- nj_map + annotate("segment", x = -74.063644, xend = -74.035, y = 40.734330, yend = 40.734330, colour = "tomato", size=.5) 
nj_map <- nj_map + annotate("point", x = -74.063644, y = 40.734330, colour = "lightblue", size = 1) 
nj_map <- nj_map + annotate("text", x = -74.01, y = 40.734330, label = "1 Journal Square", size=3, colour="gray30") 

print(nj_map)
#ggsave(nj_map, device="pdf","map1a.pdf")

nj_map <- ggplot()
nj_map <- nj_map + geom_polygon(data=nj_hf, aes(x=long, y=lat, group=group), fill=NA, color="black", size=.1)
nj_map <- nj_map + geom_polygon(data=nj_hud, aes(x=long, y=lat, group=group, fill=factor(nj_mid)), color="black", size=.5)
nj_map <- nj_map + facet_wrap(~radius)
nj_map <- nj_map + coord_map() 
#nj_map <- nj_map + scale_x_discrete()
nj_map <- nj_map + scale_fill_manual(drop=FALSE, values=c("gray", "yellow", "orange", "red", "green", "purple"), na.value="#EEEEEE", name="Unemployment rate")
nj_map <- nj_map + theme_nothing(legend=TRUE) 
nj_map <- nj_map + labs(x=NULL, y=NULL, title="1 Journal Square (2017)")
nj_map <- nj_map + theme(panel.grid.major = element_line(colour = NA))
nj_map <- nj_map + theme(text = element_text(size=15))
nj_map <- nj_map + theme(plot.title=element_text(face="bold", hjust=.4))
nj_map <- nj_map + theme(plot.subtitle=element_text(face="italic", size=9, margin=margin(l=20)))
nj_map <- nj_map + theme(plot.caption=element_text(size=12, margin=margin(t=12), color="#7a7d7e", hjust=0))
nj_map <- nj_map + theme(legend.key.size = unit(1, "cm"))
nj_map <- nj_map + annotate("segment", x = -74.063644, xend = -74.035, y = 40.734330, yend = 40.734330, colour = "tomato", size=.5) 
nj_map <- nj_map + annotate("point", x = -74.063644, y = 40.734330, colour = "lightblue", size = 1) 
nj_map <- nj_map + annotate("text", x = -74.01, y = 40.734330, label = "1 Journal Square", size=3, colour="gray30") 

print(nj_map)
#ggsave(nj_map, device="pdf","map1b.pdf")

```

For this project, officials said the qualifying rate using 2015 annual average data is 8.0 percent (1.5 times the United States not seasonally adjusted unemployment rate for the same time period).

So this project would've qualified based on any of the measures listed below.

```{r table1, fig.height=6, fig.width=9, warning=F, message=F}

sm_table <- hudson_unemployment_sm %>%
group_by(radius) %>%
summarize(average_unemployment=round(mean(un_rate, na.rm=T),2), median_unemployment=round(median(un_rate, na.rm=T),2)) %>%
arrange(average_unemployment)

kable(sm_table)


```


# 65 Bay Street (2015)

```{r map2, fig.width=9, fig.height=6, warning=F, message=F}

# GEOIDs: 34017004400, 34017004500, 34017004600, 34017005200, 34017005300, 34017005500, 34017005600, 34017005801, 34017006000, 34017006100, 34017006200, 34017006400, 34017006500, 34017006700, 34017006800, 34017007600 

#GEOIDS: 77, 78, 70, 64, 75, 74, 76
hudson_unemployment_2012 <- get_acs(geography="tract", endyear=2012, variables= c("B23025_005E", "B23025_002E"), county = "Hudson", state="NJ")
hudson_unemployment_2012$moe <- NULL

hudson_unemployment_2012 <- spread(hudson_unemployment_2012,variable, estimate )
hudson_unemployment_2012$per_un <- round(hudson_unemployment_2012[,4]/hudson_unemployment_2012[,3]*100,2)
hudson_unemployment_2012 <- filter(hudson_unemployment_2012, GEOID!="34017006900" & GEOID!="34017980100")

proj1 <- c("34017004400", "34017004500", "34017004600", "34017005200", "34017005300", "34017005500", "34017005600", "34017005801", "34017006000", "34017006100", "34017006200", "34017006400", "34017006500", "34017006700", "34017006800", "34017007600")

hudson_unemployment_2012_sm <- filter(hudson_unemployment_2012, GEOID %in% proj1)

colnames(hudson_unemployment_2012_sm) <- c("GEOID", "name", "total", "unemployed", "unemp_rate")
hudson_unemployment_2012_sm$un_rate <- hudson_unemployment_2012_sm$unemp_rate$B23025_005
hudson_unemployment_2012_sm$unemp_rate <- NULL

hudson_unemployment_2012_sm$radius <- "gerrymandered"

proj1_half <- c("34017007700", "34017007800", "34017007000", "34017006400", "34017007500", "34017007400", "34017007600")

hudson_unemployment_2012_sm_half <- filter(hudson_unemployment_2012, GEOID %in% proj1_half)
colnames(hudson_unemployment_2012_sm_half) <- c("GEOID", "name", "total", "unemployed", "unemp_rate")
hudson_unemployment_2012_sm_half$un_rate <- hudson_unemployment_2012_sm_half$unemp_rate$B23025_005
hudson_unemployment_2012_sm_half$unemp_rate <- NULL

hudson_unemployment_2012_sm_half$radius <- "adjacent tracts"

hudson_unemployment_sm <- rbind(hudson_unemployment_2012_sm, hudson_unemployment_2012_sm_half)

hudson_unemployment_sm$name <- gsub(",.*", "", hudson_unemployment_sm$name)

nj_hf <- fortify(nj_h, region="GEOID")


nj_hud <- left_join(nj_hf, hudson_unemployment_sm, by=c("id"="GEOID"))
nj_hud <- filter(nj_hud, !is.na(radius))

nj_hud$nj_mid <- cut(nj_hud$un_rate, 6)
colors <- colorRampPalette(c("white", "red"))(length(levels(nj_hud$nj_mid)))

nj_map <- ggplot()
#nj_map <- nj_map + geom_polygon(data=nj_hf, aes(x=long, y=lat, group=group), fill=NA, color="black", size=.1)
nj_map <- nj_map + geom_polygon(data=nj_hud, aes(x=long, y=lat, group=group, fill=factor(nj_mid)), color="black", size=.5)
nj_map <- nj_map + facet_wrap(~radius)
nj_map <- nj_map + coord_map() 
#nj_map <- nj_map + scale_x_discrete()
nj_map <- nj_map + scale_fill_manual(drop=FALSE, values=c("gray", "yellow", "orange", "red", "green", "purple"), na.value="#EEEEEE", name="Unemployment rate")
#nj_map <- nj_map + scale_fill_viridis(option = "heat", direction=-1, name = "Unemployment rate")
#nj_map <- nj_map + scale_color_viridis(option = "heat", direction=-1)
nj_map <- nj_map + theme_nothing(legend=TRUE) 
nj_map <- nj_map + labs(x=NULL, y=NULL, title="65 Bay Street (2015)")
nj_map <- nj_map + theme(panel.grid.major = element_line(colour = NA))
nj_map <- nj_map + theme(text = element_text(size=15))
nj_map <- nj_map + theme(plot.title=element_text(face="bold", hjust=.4))
nj_map <- nj_map + theme(plot.subtitle=element_text(face="italic", size=9, margin=margin(l=20)))
nj_map <- nj_map + theme(plot.caption=element_text(size=12, margin=margin(t=12), color="#7a7d7e", hjust=0))
nj_map <- nj_map + theme(legend.key.size = unit(1, "cm"))
nj_map <- nj_map + annotate("segment", x = -74.03577, xend = -74.005, y = 40.72, yend = 40.72, colour = "tomato", size=.5) 
nj_map <- nj_map + annotate("point", x = -74.03577, y = 40.72, colour = "lightblue", size = 1) 
nj_map <- nj_map + annotate("text", x = -73.98217, y = 40.72, label = "65 Bay Street", size=3, colour="gray30") 


print(nj_map)
#ggsave(nj_map, device="pdf", "map2a.pdf")

nj_map <- ggplot()
nj_map <- nj_map + geom_polygon(data=nj_hf, aes(x=long, y=lat, group=group), fill=NA, color="black", size=.1)
nj_map <- nj_map + geom_polygon(data=nj_hud, aes(x=long, y=lat, group=group, fill=factor(nj_mid)), color="black", size=.5)
nj_map <- nj_map + facet_wrap(~radius)
nj_map <- nj_map + coord_map() 
#nj_map <- nj_map + scale_x_discrete()
nj_map <- nj_map + scale_fill_manual(drop=FALSE, values=c("gray", "yellow", "orange", "red", "green", "purple"), na.value="#EEEEEE", name="Unemployment rate")
#nj_map <- nj_map + scale_fill_viridis(option = "heat", direction=-1, name = "Unemployment rate")
#nj_map <- nj_map + scale_color_viridis(option = "heat", direction=-1)
nj_map <- nj_map + theme_nothing(legend=TRUE) 
nj_map <- nj_map + labs(x=NULL, y=NULL, title="65 Bay Street (2015)")
nj_map <- nj_map + theme(panel.grid.major = element_line(colour = NA))
nj_map <- nj_map + theme(text = element_text(size=15))
nj_map <- nj_map + theme(plot.title=element_text(face="bold", hjust=.4))
nj_map <- nj_map + theme(plot.subtitle=element_text(face="italic", size=9, margin=margin(l=20)))
nj_map <- nj_map + theme(plot.caption=element_text(size=12, margin=margin(t=12), color="#7a7d7e", hjust=0))
nj_map <- nj_map + theme(legend.key.size = unit(1, "cm"))
nj_map <- nj_map + annotate("segment", x = -74.03577, xend = -74.005, y = 40.72, yend = 40.72, colour = "tomato", size=.5) 
nj_map <- nj_map + annotate("point", x = -74.03577, y = 40.72, colour = "lightblue", size = 1) 
nj_map <- nj_map + annotate("text", x = -73.98217, y = 40.72, label = "65 Bay Street", size=3, colour="gray30") 


print(nj_map)
#ggsave(nj_map, device="pdf", "map2b.pdf")

```

For this project, officials said the qualifying rate using 2015 annual average data is 9.3 percent (1.5 times the United States not seasonally adjusted unemployment rate for the same time period).

So this project would've qualified based on only the gerrymandered version since unemployment was too low in the other ways of grouping the census tracts.


```{r table2, fig.height=6, fig.width=9, warning=F, message=F}

sm_table <- hudson_unemployment_sm %>%
  group_by(radius) %>%
  summarize(average_unemployment=round(mean(un_rate, na.rm=T),2), median_unemployment=round(median(un_rate, na.rm=T),2)) %>%
  arrange(average_unemployment)

kable(sm_table)


```

